### 2021-04-19 doc/VerticalArrangementsSheuTaragin.tex
added text on bargaining power. Tweaked figures

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Vertically Integrated Incumbent Firms}
Here, we investigate how the presence of one or more incumbent vertically integrated firms affects merger outcomes when firms either have constant or linear marginal costs. In particular, we examine merger outcomes when a vertically integrated merges with either an unintegrated upstream or unintegrated downstream firm, or another integrated firm. We also examine merger outcomes for these integrated mergers as the number of integrated non-merging firms increases.

Figure \ref{fig:CVvertincumbBW} depicts box and whisker plots summarizing consumer harm (top row) and total harm (bottom row) as the number of incumbent integrated firms increases from 0 to 6 firms. The box and whisker  plots when the number of  incumbent integrated firms equals 0 correspond to the results depicted in Figure 1 of \citeN{ST2017}. For downstream and upstream  mergers, the plots when the number of  incumbent integrated firms equals 1 depict the outcome of a merger between an integrated firm and an unintegrated rival, and all 3rd parties are unintegrated. By contrast, for vertical mergers, the plots when the number of  incumbent integrated firms equals 1 depict the outcome from an unintegrated wholesaler merging with an unintegrated retailer when a single 3rd party is integrated. Likewise, for integrated mergers, the plots when the number of  incumbent integrated firms equals 2 depict a merger between two integrated firms, and all rivals are unintegrated. The blue box and whisker plots (left) assume that firms face constant marginal costs, while the orange box and whisker plots (right) assume that firms face linear marginal costs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Bargaining Power}

@@

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.9]{../output/CVvertincumbBW.png}
\caption{The figure displays box and whisker plots summarizing the extent to which mergers affect consumer and total surplus as the number of vertically integrated firms present in a market change. Each blue box depicts the effects assuming that firms face constant marginal costs, while each orange box depicts the effects assuming that firms face linear marginal costs. Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median. }
\label{fig:CVvertincumbBW}
\end{sidewaysfigure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Vertically Integrated Incumbent Firms}
Here, we examine merger outcomes when a vertically integrated merges with either an unintegrated upstream or unintegrated downstream firm, or another integrated firm. We also examine merger outcomes for these integrated mergers as the number of integrated non-merging firms increases.

Figure \ref{fig:CVvertincumbBW} depicts box and whisker plots summarizing consumer harm (blue,left) and total harm (orange, right) as the number of incumbent integrated firms increases from 0 to 6 firms. The box and whisker  plots when the number of  incumbent integrated firms equals 0 correspond to the results depicted in Figure 1 of \citeN{ST2017}. For downstream and upstream  mergers, the plots when the number of  incumbent integrated firms equals 1 depict the outcome of a merger between an integrated firm and an unintegrated rival, and all 3rd parties are unintegrated. By contrast, for vertical mergers, the plots when the number of  incumbent integrated firms equals 1 depict the outcome from an unintegrated wholesaler merging with an unintegrated retailer when a single 3rd party is integrated. Likewise, for integrated mergers, the plots when the number of  incumbent integrated firms equals 2 depict a merger between two integrated firms, and all rivals are unintegrated.

Starting with downstream mergers (leftmost column), absent incumbent vertical integration (0 integrated firms), not only are downstream mergers rarely beneficial, but median consumer harm  from downstream mergers is about 9\% of pre-merger revenues while total harm is about 5\% of pre-merger revenues. A merger between an incumbent integrated firm and an unintegrated retailer when all other market participants are unintegrated (1 integrated firm)  increases the range of outcomes while also increasing median harm: consumer welfare is positive in more than 13\% of mergers and total welfare is postive in 3.4\%, but median consumer harm increases to 14\% and median total harm increases to 7.6\% of pre-merger revenues. Increasing the number of integrated 3rd parties to 1 (2 integrated firms) shrinks the inter-quartile range but has little impact on median harm. Adding additional integrated 3rd parties (3 integrated firms) further decreases the range of outcomes while also decreasing median harm. With 3 integrated 3rd parties (4 integrated firms), a downstream merger between an integrated and unintegrated firms benefits consumers in about 20\% of markets and shrinks median consumer harm to about 6\% of pre-merger revenues. By contrast, with 3 integrated 3rd parties, the inter-quartile range shrinks, and median total harm falls to about 3.7\% of pre-merger revenues.

Like downstream mergers, not only are upstream mergers absent incumbent vertical integration rarely beneficial, but median consumer harm  from upstream mergers is about 6\% of pre-merger revenues  and total harm is about 3\% of pre-merger revenues. A merger between an incumbent integrated firm and an unintegrated retailer when all other market participants are unintegrated (1 integrated firm)  increases the proportion of beneficial mergers but does little to change median harm: consumer welfare is positive in less than 8\% of mergers and total welfare is positive in 2.3\% of mergers, but median consumer harm increases to 7\% and median total harm increases to 4.8\% of pre-merger revenues. Increasing the number of integrated 3rd parties to 1 (2 integrated firms) shrinks the inter-quartile range but has little impact on median harm. Adding additional integrated 3rd parties (3,4 integrated firms) further decreases the range of outcomes while leaving median harm largely unchanged.

Turning to vertical mergers, absent incumbent integration, about 22\% of vertical mergers benefit consumers and about 14\% of vertical mergers increase total surplus.  The presence of a single 3rd party integrated firm (1 integrated firm) decreases the percentage of mergers that benefit consumers to 14.5\% and the percentage of mergers that increase  total surplus to 9.6\%, but leads median harm largely unchanged. Interestingly, while adding a second integrated incumbent to the market also lowers the percentage of beneficial mergers,adding a third integrated firm increases the percentage of mergers that benefits consumers to about 15\%, and the percentage of mergers that increase total surplus to 6.9\%. Adding a fourth integrated incumbent has a similar effect to adding a third incumbent.

Finally, absent the presence of rival incumbent integrated firms, mergers between two integrated firms harm consumers in about 96\% of simulated markets and lowers total surplus in more than 92\% of simulated markets. while adding one additional rival incumbent integrated firm (3 integrated firms) has little effect on either the distribution of consumer or total surplus, additional integrated rivals eliminate the most harmful integrated mergers while also increasing median consumer harm by as much as 13.3\% (5 integrated firms) and median total harm by as much as 8.4\%.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Bargaining Power}
\citeN{ST2017} show using numerical simulation that downstream and vertical mergers where wholesalers have relatively more bargaining power are less harmful than mergers where retailers have relatively more bargaining power. Here, we show that this result continues to hold when there are rival incumbent integrated firms. We also show that this relationship holds for integrated mergers as well as upstream mergers when one of the merging parties is vertically integrated.

Figures \ref{fig:CVbargupBW}-\ref{fig:CVbargbothBW}  depicts box and whisker plots summarizing the consumer, retailer, wholesaler and total welfare effects for each merger as the number of incumbent integrated firms increases. For upstream, downstream and vertical mergers, the figures depict box and whisker plots for 0,1 and 4 incumbent vertically integrated firms.  For downstream and upstream  mergers (Figures \ref{fig:CVbargupBW} and \ref{fig:CVbargdownBW}), the plots when the number of  incumbent integrated firms equals 1 depict the outcome of a merger between an integrated firm and an unintegrated rival, and all 3rd parties are unintegrated. By contrast, for vertical mergers ((Figures \ref{fig:CVbargvertBW}), the plots when the number of  incumbent integrated firms equals 1 depict the outcome from an unintegrated wholesaler merging with an unintegrated retailer when a single 3rd party is integrated. Finally, Figure \ref{fig:CVbargbothBW} depicts box and whisker plots for integrated mergers with 2,4, or 6 integrated firms.  When the number of  incumbent integrated firms equals 2, Figure \ref{fig:CVbargbothBW} depicts a merger between two integrated firms, with all rivals are unintegrated. Plots with either 4 or 6 integrated firms depict mergers between integrated incumbents when there are either 2 or 4 rival integrated firms in the market.

@@

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.9]{../output/CVvertincumbBW.png}
\caption{The figure displays box and whisker plots summarizing the extent to which mergers affect consumer (blue,left) and total (orange,right) surplus as the number of vertically integrated firms present in a market change.  Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median. }
\label{fig:CVvertincumbBW}
\end{sidewaysfigure}
